---
title: "Causal Inference with Inteference"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Causal Inference with Inteference}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# RCT2 
This package provides various statistical methods for designing and analyzing two-stage randomized controlled trials. Two-stage randomized controlled trials can be used to estimate spillover effects as well as direct treatment effects.

# Overview
There are three functions in this package:

1. CADErand: computes the point estimates and variance estimates of the complier average direct effect (CADE) and the complier average spillover effect (CASE). The estimators calculated using this function are either individual weighted or cluster-weighted. The point estimates and variances of ITT effects are also included.

2. CADEreg: computes the point estimates of the complier average direct effect (CADE) and four different variance estimates: the HC2 variance, the cluster-robust variance, the cluster-robust HC2 variance and the variance proposed in the reference.

3. CADEparamreg: computes the point estimates of the complier average direct effect (CADE) and the complier average spillover effect (CASE) following the model-based approach presented in the appendix.

# Functions
Before we begin, lets load the library and our example data set into R.

```{r, include=FALSE}
devtools::load_all(".")
library(RCT2)
data <- RCT2::india
```


## CADErand
To run the CADErand command, simply type in the following:
```{r}
data$id <- factor(data$id)
rand <- CADErand(data)
printCADE(rand)
```

Then, you will be able to access the output of the CADE randomization function including the CADE, CASE, ITT effects, and the variances of all of those things.

```{r}
rand$CADE1
rand$CADE0
```

## CADEreg
In order to analyze our data using a regression based method, we use the CADEreg function.
```{r}
reg <- CADEreg(data, ci.level = 0.90)
printCADE(reg)
```

This gives us the point estimates of CADE1 and CADE0 and their confidence intervals, and various types of variances for the CADE1 and CADE0. We can again access these by using the dollar sign notation. Note that we can use the parameter \code{ci.level} to specify the confidence interval level (i.e. 95\%, 90\%).
```{r}
reg$CADE1
reg$CADE0
```


## CADEparamreg
CADEparamreg offers a regression-based method for the computing the ITT effects and the average direct effects and spillover effects.
```{r}
paramreg <- CADEparamreg(data, assign.prob = 0.8, ci.level = 0.95)
printCADE(paramreg)[[1]]
```

Note how we use \code{assign.prob} to specify the assignment probability to the different assignment mechanisms. We also use \code{ci.level} again to specify the confidence intervals.
